---
title: "Military Surplus import/clean"
output: html_notebook
---

## Goals of this notebook

- Download the data
- Import the data
- Clean datatypes
- Remove unnecessary columns
- Create a total_value column
- Filter to Texas agencies
- Filter the date range (since Jan. 1 2010)
- Export the cleaned data

## Project Setup

Setup our project environment

```{r setup}
library(tidyverse)
```

## Download Defense Data

The Defense Logistics Agency transfers surplus military equipment to local law enforcement through its [Law Enforcement Support Office](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/PublicInformation/). You can find more information [about the program here](https://www.dla.mil/DispositionServices/Offers/Reutilization/LawEnforcement/ProgramFAQs/)

```{r download_data} 
download.file("https://github.com/utdata/rwd-r-leso/blob/main/data-processed/leso.csv?raw=true", "data-raw/leso.csv")
```

## Read data

Read the csv file imported

```{r read}

leso <- read_csv("data-raw/leso.csv") # assign data to new tibble object

leso # print tibble as table
```

```{r glimpse}
#glimpse tibble to see columns
leso %>% glimpse()
```

## Clean up our columns

Use select to get rid of the columns we don't need!

```{r select_columns}
leso_tight <- leso %>% 
  select(
    -sheet, 
    -nsn,
    -starts_with("demil")
  )

leso_tight %>% glimpse()
```

## Create a total value column

Use mutate to get the total value of items in each row

```{r mutate}
# create new tibble with a total value column
 leso_total <- leso_tight %>% 
  mutate(total_value = quantity * acquisition_value) 

# glimpse our new tibble
leso_total %>% glimpse()
```

## Filter for Texas data

Use filter to find only the data pertinent to Texas police departments since January 01, 2010

```{r filter_data}
# filter data for only Texas departments after Jan. 1, 2010
leso_filtered <- leso_total %>% 
  filter(
    state == "TX",
    ship_date >= "2010-01-01"
  ) 

leso_filtered %>% glimpse()
```

## Export data

Now that we've gotten rid of everything we don't need, let's export shall we?

```{r export_filtered}
# filter our texas-only data into new file
leso_filtered %>% write_rds("data-processed/01-leso-tx.rds")
```
