---
title: "Military surplus analysis"
output:
  html_document:
    df_print: paged
---

## Setup

Get the tidyverse library all up in this file!

```{r setup}
library(tidyverse)
```

## Import Data

Import our cleaned and filtered data into this notebook

```{r import_data}
#import from processed folder
tx <- read_rds("data-processed/01-leso-tx.rds")

# glimpse our data
tx %>% glimpse()
```

## Agency totals

For each agency in Texas, find the summed **quantity** and summed **total value** of the equipment they received.

```{r summed_quant_and_totals}
tx_quant_totals <- tx %>% 
  group_by(agency_name) %>% # group by agency
  summarize(
    summed_quantity = sum(quantity),
    summed_total_val = sum(total_value)
  ) %>% # sum total quantity and total values
  arrange(summed_total_val %>% desc()) # Arrange by dept with most total val

#peek at results
tx_quant_totals
```

#### Texas total gifts

Take a closer look at how much value went to Texas overall

```{r texas_totals_val}
tx %>% 
  summarize(
    summed_quantity = sum(quantity),
    summed_total_val = sum(total_value)
  ) 
```


## Look at local cities

Filter summed quantity and totals data for only specific local departments

```{r local_filter}
# create list of reusable local dept.'s
local_agencies <- c(
  "AUSTIN POLICE DEPT",
  "SAN MARCOS POLICE DEPT",
  "TRAVIS COUNTY SHERIFFS OFFICE",
  "UNIV OF TEXAS SYSTEM POLICE HI_ED",
  "WILLIAMSON COUNTY SHERIFF'S OFFICE"
)

# filter quant/total data for only local agencies
tx_quant_totals %>% 
  filter(agency_name %in% local_agencies)
```
## Breakdown local agency items

Finding the sums of each different item each local agency has received since 2010

### Austin Police Department items

What did our city buy the last 10 years?

**Austin Population**: 961,855 (2020 Census)


```{r items_austin_10}
tx %>% 
  filter(
    agency_name == "AUSTIN POLICE DEPT" # get only Austin entries
  ) %>% 
  group_by(item_name) %>% # group same items together
  summarize(
    summed_quantity = sum(quantity), #sum all quantities
    summed_total_val = sum(total_value) # sum all totals
  ) %>% 
  arrange(summed_total_val %>% desc()) #arrange by most expensive
```

### Last 5 years (since Jan. 01, 2015)

Let's take a closer look

```{r items_austin_five}
tx %>% 
  filter(
    agency_name == "AUSTIN POLICE DEPT", # get only Austin entries
    ship_date >= "2015-01-01"
  ) %>% 
  group_by(item_name) %>% # group same items together
  summarize(
    summed_quantity = sum(quantity), #sum all quantities
    summed_total_val = sum(total_value) # sum all totals
  ) %>% 
  arrange(summed_total_val %>% desc()) #arrange by most expensive
```
```{r austin_totals_five}
tx %>% 
  filter(
    agency_name == "AUSTIN POLICE DEPT", # get only Austin entries
    ship_date >= "2015-01-01"
  ) %>% 
  summarize(
    summed_quantity = sum(quantity),
    summed_total_val = sum(total_value)
  ) 
```

```{r austin_totals_ten}
tx %>% 
  filter(
    agency_name == "AUSTIN POLICE DEPT", # get only Austin entries
  ) %>% 
  summarize(
    summed_quantity = sum(quantity),
    summed_total_val = sum(total_value)
  ) 
```


### San Marcos Police Dept. items

Ok, let's take a look at San Marcos

**San Marcos Population**: 67,553 (2020 Census)

```{r items_san_marcos}
tx %>% 
  filter(
    agency_name == "SAN MARCOS POLICE DEPT" # get only San Marcos entries
  ) %>% 
  group_by(item_name) %>% # group same items together
  summarize(
    summed_quantity = sum(quantity), #sum all quantities
    summed_total_val = sum(total_value) # sum all totals
  ) %>% 
  arrange(summed_total_val %>% desc()) #arrange by most expensive
```

### Travis Co. Sheriff's Office items

Ok, let's take a look at Travis County

**Travis County Population**: 1,290,188 (2020 Census)

```{r items_travis_co}
tx %>% 
  filter(
    agency_name == "TRAVIS COUNTY SHERIFFS OFFICE" # get only Travis Co. Sheriff entries
  ) %>% 
  group_by(item_name) %>% # group same items together
  summarize(
    summed_quantity = sum(quantity), #sum all quantities
    summed_total_val = sum(total_value) # sum all totals
  ) %>% 
  arrange(summed_total_val %>% desc()) #arrange by most expensive
```

### University of Texas Police

Ok, now let's look at UT

**Travis County Population**: 1,290,188 (2020 Census)

```{r items_ut_austin}
tx %>% 
  filter(
    agency_name == "UNIV OF TEXAS SYSTEM POLICE HI_ED" # get only Travis Co. Sheriff entries
  ) %>% 
  group_by(item_name) %>% # group same items together
  summarize(
    summed_quantity = sum(quantity), #sum all quantities
    summed_total_val = sum(total_value) # sum all totals
  ) %>% 
  arrange(summed_total_val %>% desc()) #arrange by most expensive
```

### Williamson County Sheriff's Office items

Lastly, let's take a look at Williamson

**Williamson County Population**: 609,017 (2020 Census)

```{r items_williamson_co}
tx %>% 
  filter(
    agency_name == "WILLIAMSON COUNTY SHERIFF'S OFFICE" # get only Williamson Co. Sheriff entries
  ) %>% 
  group_by(item_name) %>% # group same items together
  summarize(
    summed_quantity = sum(quantity), #sum all quantities
    summed_total_val = sum(total_value) # sum all totals
  ) %>% 
  arrange(summed_total_val %>% desc()) #arrange by most expensive
```
## Some notable equipment

- **MINE RESISTANT VEHICLE**: Mine-Resistant, Ambush-Protected vehicle developed to withstand impact from roadside bombs in Iraq and Afghanistan. Mostly used in parades, but have been deployed in standoffs with active shooters. (San Marcos: 1, Travis County: 1, )
- **UNMANNED VEHICLE,GROUND**: UGVs; Robotics vehicle used for many applications militarily like explosive ordinance disposal (EOD) such as landmines, loading heavy items, and repairing ground conditions under enemy fire; Civilian uses include urban search and rescue, fire fighting, and nuclear response. (San Marcos: 3)
- **510 PackBot**:  tactical mobile robot developed by Endeavor Robotics; used for high-threat scenarios such as surveillance, chemical and biological detection, and bomb disposal. (Travis County: 1, Williamson County: 2, San Marcos: 1, Austin: 4)
